---
title: "R Notebook"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
---


# Load Libraries

```{r}
library(data.table)
library(factoextra)
```

# Load Data

```{r}
full_data <- fread("../input_data.csv")
X_data <- full_data[,-c('y')]
y_data <- full_data[, y]
```


# Inspect Dependent Variable

```{r}
hist(y_data)
```


The min value of the dependent variable is: `r min(y_data)`.  
The max value of the dependent variable is: `r max(y_data)`.  
The number if missing dependent values is: `r sum(is.na(y_data))`.  


# Inspect Independent Variable

Determine if any fields in the data set are non numeric.

```{r}
# Converting non-numeric values to numeric will create an NA
which(is.na(as.numeric(as.character(X_data[[1]]))))
```


Determine which columns contain nulls and what percent of values are null.

```{r}
# Calculate percent null using mean NA by column
nullCols <- sort(colMeans(is.na(X_data)))
nullCols[nullCols > 0]
```


Determine which columns have zero variance.

```{r}
# Calculate unique counts by column
uniqueCount <- sort(rapply(X_data,function(x)length(unique(x))))
uniqueCount[uniqueCount == 1]
```


Remove zero variance columns.

```{r}
# Remove four zero variance columns
X_data <- X_data[, -c('x067', 'x094', 'x095', 'x096')]
```


Impute missing data using the median of each column.

```{r}
# Impute median for missing values
replmed <- function(x) replace(x, is.na(x), median(x, na.rm = TRUE))
X_data <- replace(X_data, TRUE, lapply(X_data, replmed))
```


Determine which columns are perfectly correlated with a prior column.

```{r}
corX <- (cor(X_data,X_data)==1)
corX[lower.tri(corX, diag=TRUE)] <- FALSE

## Extract the names of the redundant columns
colnames(corX)[colSums(corX)>0]
```


Remove perfectly correlated columns

```{r}
# Remove one perfectly correlated column
X_data <- X_data[, -c('x276')]
```


Display scree plot showing variance captured by top principal components

```{r}
#pca <- princomp(X_data, cor=TRUE, scores=FALSE)
pca <- prcomp(X_data, center=TRUE, scale=TRUE)
fviz_eig(pca)
plot(summary(pca)$importance[3,], xlab="Principal Components", ylab="Cumulative Variance Explained", main="Principal Component Summary")
```

